Classification of ordinal data in deep learning: an experimental study

### 2019-01-11 paper/ieee.tex
Correcciones paper

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
% \usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../pdf/}{../jpeg/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
% or other class option (dvipsone, dvipdf, if not using dvips). graphicx
% will default to the driver specified in the system graphics.cfg if no
% driver is specified.
% \usepackage[dvips]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../eps/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

@@

	Softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems but not the best when working with ordinal data. In this paper, the Proportional Odds Model (POM)~\cite{agresti2010analysis} will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimization algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust results. An approximated ANOVA III test will be performed because of the limitations of the computational time required to run a higher number of executions.

@@

	\section{Experiments}
	\label{sect:experiments}
	\subsection{Data}
	We make use of two ordinal datasets appropriate for deep neural networks:

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try and predict from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras, illumination conditions and resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition that was hosted on the Kaggle platform. Also, this dataset was used in later works~\cite{de2018weighted, nebot2016diabetic} and ordinal techniques (such as an ordinal cost function) were applied in order to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels. Data augmentation techniques are applied to achieve a higher number of samples.

@@

	\subsection{The model}
	A convolutional neural network (CNN) has been used for both datasets. The architecture of this CNN is presented in the Table \ref{table:CNNArchitecture}.

@@

	Every convolutional layer is followed by an ELU activation layer~\cite{clevert2015fast} and a batch normalization~\cite{ioffe2015batch}. At the output, a Proportional Odds Model is used with different link functions~\cite{agresti2010analysis}. The logit link function is commonly used within POM. In this paper, we are comparing other link functions like probit or complementary log-log with the logit link. These three types of links are explained below.

	\begin{itemize}
		\item \textit{Logit}. Logit link function is the most widely used function for Proportional Odds Models. These kind of models are also called Cummulative Logit Models. The logit link is shown in Eq. \ref{eq:logit}.
		\begin{equation}
		logit[P(Y_i \le j)] = \alpha_j + \beta'x_i,\quad j = 1, ..., c - 1
		\label{eq:logit}
		\end{equation}
		\item \textit{Probit}. Probit link function is the inverse of the standard normal cummulative distribution function (cdf). Its expression is shown in Eq. \ref{eq:probit}.	
		\begin{equation}
		\Phi^{-1}[P(Y \le j)] = \alpha_j + \beta'x,\quad j = 1, ..., c - 1
		\label{eq:probit}
		\end{equation}
		\item \textit{Complementary log-log}. Unlike logit and probit, complementary log-log function is not symmetric. With a continuous predictor $x$, for example, $P(Y \le j)$ approaches 0 at a different rate than it approaches 1. Complementary log-log expression is shown in Eq. \ref{eq:cloglog}.
		\begin{equation}
		\small
		log[-log[1 - P(Y \le j)]] = \alpha_j + \beta'x,\quad j = 1, ..., c - 1
		\label{eq:cloglog}
		\end{equation}
	\end{itemize}

@@

	Quadratic Weighted Kappa Loss is considered as loss function for this optimizer as it gives better performance for ordinal classification problems.

@@

	The model is evaluated with Quadratic Weighted Kappa metric (QWK)~\cite{ben2008comparison}. This evaluation measure gives a higher weight to the errors that are further from the correct class.

@@

	\subsection{Parameters}
	Three different parameters have been considered: learning rate, batch size and link function for the final output layer.

@@

	\begin{table}[!t]
		\caption{Diabetic Retinopathy results. BS stands for Batch Size, LR for Learning Rate and LF for link function.}
		\label{table:DRresults}
		\footnotesize
		\centering
		\begin{tabular}{ccc|cc}
			BS & LR & LF & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			$5$ & $10^{-2}$ & Logit & $0.44888$ & $0.4163$\\
			& & Probit & $0.46724$ & $0.45972$\\
			& & c log-log & $0.42854$ & $0.41448$\\
			& $10^{-3}$ & Logit & $0.56496$ & $0.55356$\\
			& & Probit & $0.57084$ & $0.56392$\\
			& & c log-log & $0.54796$ & $0.53436$\\
			& $10^{-4}$ & Logit & $0.52884$ & $0.52032$\\
			& & Probit & $0.53802$ & $0.52302$\\
			& & c log-log & $0.53824$ & $0.51974$\\
			$10$ & $10^{-2}$ & Logit & $0.5497$ & $0.53142$\\
			& & Probit & $0.52124$ & $0.50806$\\
			& & c log-log & $0.4276$ & $0.4227$\\
			& $10^{-3}$ & Logit & $0.58036$ & $0.57686$\\
			& & Probit & $0.56536$ & $0.5581$\\
			& & c log-log & $\textbf{0.5883}$ & $\textbf{0.5815}$\\
			& $10^{-4}$ & Logit & $0.53158$ & $0.5385$\\
			& & Probit & $0.5456$ & $0.54082$\\
			& & c log-log & $0.53928$ & $0.53742$\\
			$15$ & $10^{-2}$ & Logit & $0.55828$ & $0.55076$\\
			& & Probit & $0.54112$ & $0.5343$\\
			& & c log-log & $0.56676$ & $0.56428$\\
			& $10^{-3}$ & Logit & $0.55748$ & $0.55106$\\
			& & Probit & $0.58182$ & $0.57894$\\
			& & c log-log & $0.5634$ & $0.55938$\\
			& $10^{-4}$ & Logit & $0.54686$ & $0.54342$\\
			& & Probit & $0.53258$ & $0.53278$\\
			& & c log-log & $0.539$ & $0.53828$
		\end{tabular}
	\end{table}

@@

	\begin{table}[ht]
		\caption{Adience results. BS stands for Batch Size, LR for Learning Rate and LF for link function.}
		\label{table:Adienceresults}
		\footnotesize
		\centering
		\begin{tabular}{ccc|cc}
			BS & LR & LF & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			$5$ & $10^{-2}$ & Logit & $0.72072$ & $0.68408$\\
			& & Probit & $0.77608$ & $0.72574$\\
			& & c log-log & $0.7543$ & $0.70742$\\
			& $10^{-3}$ & Logit & $0.7146$ & $0.67868$\\
			& & Probit & $0.81906$ & $0.77154$\\
			& & c log-log & $0.57196$ & $0.55608$\\
			& $10^{-4}$ & Logit & $0.54432$ & $0.51934$\\
			& & Probit & $0.5436$ & $0.52422$\\
			& & c log-log & $0.54802$ & $0.52484$\\
			$10$ & $10^{-2}$ & Logit & $0.78872$ & $0.72994$\\
			& & Probit & $0.79332$ & $0.7475$\\
			& & c log-log & $0.78004$ & $0.7326$\\
			& $10^{-3}$ & Logit & $0.49686$ & $0.47112$\\
			& & Probit & $0.59778$ & $0.58476$\\
			& & c log-log & $0.52016$ & $0.49968$\\
			& $10^{-4}$ & Logit & $0.60972$ & $0.56814$\\
			& & Probit & $0.54638$ & $0.52632$\\
			& & c log-log & $0.75428$ & $0.52094$\\
			$15$ & $10^{-2}$ & Logit & $0.79104$ & $0.74388$\\
			& & Probit & $\textbf{0.84254}$ & $\textbf{0.78666}$\\
			& & c log-log & $0.78802$ & $0.74624$\\
			& $10^{-3}$ & Logit & $0.56946$ & $0.52978$\\
			& & Probit & $0.70262$ & $0.65416$\\
			& & c log-log & $0.49242$ & $0.4732$\\
			& $10^{-4}$ & Logit & $0.55202$ & $0.52484$\\
			& & Probit & $0.54288$ & $0.5174$\\
			& & c log-log & $0.53826$ & $0.52542$
		\end{tabular}

@@

	\subsection{Statistical analysis}
	\label{sect:statisticalanalysis}
	In this subsection, a statistical analysis will be performed in order to obtain robust conclusions from the results presented in this section.

@@

	We denote by $W_{i,j,k,l}(i=1, ..., 3; j = 1, ..., 3; k = 1, ..., 3)$ the value observed when the first factor is at the i-th level, the second at the j-th level and the third at k-th level. We assume that the two factors do not act independently and therefore there exists an interaction between them. In this case, the observations fit Eq. \ref{eq:anova3}.

	\begin{equation}
	\begin{aligned}
	\footnotesize
	W_{i,j,k,l} = & \mu + L_i + P_j + B_k + LP_{i,j} + LB_{i,k} \\& + PB_{j,k} + LPB_{i,j,k} + \epsilon_{i,j,k,l}
	\label{eq:anova3}
	\end{aligned}
	\end{equation}

@@

	\begin{table}[ht]
		\caption{ANOVA III for the analysis of the main factors in the design of a Convolutional Ordinal Neural Network for the Retinopathy dataset.}
		\label{table:ANOVARetinopathy}
		\centering
		\small
		\begin{tabular}{l|ccccc}
			\multicolumn{6}{c}{Response variable WK}\\\hline
			Source & S.S. & D.F. & M.S. & F-ratio & Sig.\\\hline
			Model & $37.860$ & $9$ & $4.207$ & $1562.840$ & $.000$\\
			$L$ factor & $.057$ & $2$ & $.029$ & $10.646$ & $.000$\\
			$P$ factor & $.121$ & $2$ & $.060$ & $22.468$ & $.000$\\
			$LP$ factors & $.057$ & $4$ & $.014$ & $5.261$ & $.001$\\
			Error & $.339$ & $126$ & $.003$ & & \\
			Total & $38.199$ & $135$ & & & 
		\end{tabular}
	\end{table}

@@

	\begin{table}[ht]
		\caption{ANOVA III for the analysis of the main factors in the design of a Convolutional Ordinal Neural Network for the Adience dataset.}
		\label{table:ANOVAAdience}
		\centering
		\small
		\begin{tabular}{l|ccccc}
			\multicolumn{6}{c}{Response variable WK}\\\hline
			Source & S.S. & D.F. & M.S. & F-ratio & Sig.\\\hline
			Model & $52.328$ & $15$ & $3.489$ & $1426.170$ & $.000$\\
			$L$ factor & $.027$ & $2$ & $.014$ & $5.582$ & $.005$\\
			$P$ factor & $1.031$ & $2$ & $.516$ & $210.809$ & $.000$\\
			$B$ factor & $.089$ & $2$ & $.045$ & $18.253$ & $.000$\\
			$LP$ factors & $.183$ & $4$ & $.046$ & $18.695$ & $.000$\\
			$PB$ factors & $.124$ & $4$ & $.031$ & $12.695$ & $.000$\\
			Error & $.294$ & $120$ & $.002$ & & \\
			Total & $52.622$ & $135$ & $3.489$ & & 
		\end{tabular}
	\end{table}

@@

	\begin{table}[ht]
		\caption{Nominal method results. BS stands for Batch Size and LR for Learning Rate.}
		\label{table:NominalResults}
		\footnotesize
		\centering
		\begin{tabular}{lcc|cc}
			Dataset & BS & LR & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			Retinopathy & $5$ & $10^{-2}$ & $0.2210$ & $0.2050$\\
			Retinopathy & $5$ & $10^{-3}$ & $0.2807$ & $0.2654$\\
			Retinopathy & $5$ & $10^{-4}$ & $0.4580$ & $0.4449$\\

			Retinopathy & $10$ & $10^{-2}$ & $0.2972$ & $0.2997$\\
			Retinopathy & $10$ & $10^{-3}$ & $0.3972$ & $0.3989$\\
			Retinopathy & $10$ & $10^{-4}$ & $\textbf{0.4972}$ & $\textbf{0.4967}$\\

			Retinopathy & $15$ & $10^{-2}$ & $0.3692$ & $0.3676$\\
			Retinopathy & $15$ & $10^{-3}$ & $0.4107$ & $0.4162$\\
			Retinopathy & $15$ & $10^{-4}$ & $0.4929$ & $0.4859$\\
			\hline
			Adience & $5$ & $10^{-2}$ & $0.0605$ & $0.0769$\\
			Adience & $5$ & $10^{-3}$ & $0.7994$ & $0.7313$\\
			Adience & $5$ & $10^{-4}$ & $0.8009$ & $0.7383$\\

			Adience & $10$ & $10^{-2}$ & $0.0559$ & $0.0576$\\
			Adience & $10$ & $10^{-3}$ & $0.8239$ & $0.7436$\\
			Adience & $10$ & $10^{-4}$ & $0.7938$ & $0.7218$\\

			Adience & $15$ & $10^{-2}$ & $0.0593$ & $0.0713$\\
			Adience & $15$ & $10^{-3}$ & $\textbf{0.8309}$ & $\textbf{0.7575}$\\
			Adience & $15$ & $10^{-4}$ & $0.7867$ & $0.7129$\\
		\end{tabular}
	\end{table}

>>>>

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../pdf/}{../jpeg/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
% or other class option (dvipsone, dvipdf, if not using dvips). graphicx
% will default to the driver specified in the system graphics.cfg if no
% driver is specified.
% \usepackage[dvips]{graphicx}
% declare the path(s) where your graphic files are
% \graphicspath{{../eps/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
% \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\usepackage{dsfont}
\usepackage{amssymb}

@@

	Softmax function is widely used for the output layer in neural networks for classification tasks. It is a simple and efficient function for multi-class problems but not the best when working with ordinal data. In this paper, the Proportional Odds Model (POM)~\cite{gutierrez2016ordinal, agresti2010analysis} will be used instead of the softmax function. Different link functions will be explored to compare their performance. Also, the influence of other parameters like the learning rate of the optimization algorithm and the batch size will be studied. We will contrast the results obtained with statistical analysis to provide more robust results. An approximated ANOVA III test will be performed because of the limitations of the computational time required to run a higher number of executions.

@@

	\section{Ordinal problem definition}
	\label{sect:ordinalproblem}
	The ordinal problem consists of predicting the label y of an input vector x, where $x \in \mathcal{X} \subseteq \mathds{R}^K$ and $y \in \mathcal{Y}~=~\{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_Q\}$, i.e. x is in a K-dimensional input space, and $y$ is in a label space of $Q$ different labels. The objective of the ordinal problem is to find a function $r : \mathcal{X} \rightarrow \mathcal{Y}$ to predict the labels or categories of new patterns, given a training set of $N$ points, $D = \{(x_i, y_i), i = 1, ..., N\}$. A natural label ordering exists for ordinal problems: $\mathcal{C}_1 \prec \mathcal{C}_2 \prec ... \prec \mathcal{C}_Q$. The assumption of order between class labels makes that two different elements of $\mathcal{Y}$ can always be compared by using the relation $\prec$, which is not possible under the nominal classification setting. In regression (where $y \in \mathds{R}$), real values in $\mathds{R}$ can be ordered by the standard $<$ operator, but labels in ordinal regression ($y \in \mathcal{Y}$) do not carry metric information, so the category serves as a qualitative indication of the pattern rather than a quantitative one.

	\section{Proportional Odds Model}
	\label{sect:pom}
	Arising from a statistical background, the Proportional Odds Model (POM) is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}, dated back to 1980. It is a member of a wider family of models recognised as Cumulative Link Models (CLMs)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec Cj |x)$ are estimated, which can be directly related to standard probabilities:
	\[
		P(y \preceq \mathcal{C}_q | x) = P(y = \mathcal{C}_1 | x) + ... + P(y = \mathcal{C}_q | x),
	\]
	\[
		P(y = \mathcal{C}_q | x) = P(y \preceq \mathcal{C}_q | x) - P(y \preceq \mathcal{C}_{q-1} | x),
	\]
	with $q = 2, ..., Q$, and considering that $P(y = \mathcal{C}_1 | x) = P(y \preceq \mathcal{C}_1 | x)$ and $P(y \preceq \mathcal{C}_Q | x) = 1$.

	The model in inspired by the notion of latent variable, considering a linear transformation $f(x) = w^Tx$ of the input variables as the one-dimensional mapping. The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\[
		g^{-1}(P(y \preceq \mathcal{C}_q | x)) = b_q - w^Tx, \quad 1, ..., Q-1,
	\]
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y* = f(x)* = w^Tx + \epsilon$, where $\epsilon$ is the random error component. The most common choice for the distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is observed if and only if $f(x) \in [b_{q-1}, b_q]$, where the function $f$ and $b = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. it is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(x), x \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each region corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | x)$ increases with $q$ \cite{mccullagh1980regression}.
	In this work, we use this ordinal model with different link functions, including logit, probit and complementary log-log. These functions will be detailed in Section \ref{sect:model}.
	\section{Experiments}
	\label{sect:experiments}
	\subsection{Data}
	We make use of two ordinal datasets appropriate for deep neural networks:

	\begin{itemize}
		\item \textit{Diabetic Retinopathy (DR)}\footnote{https://www.kaggle.com/c/diabetic-retinopathy-detection/data}. DR is a dataset consisting of extremely high-resolution fundus image data. The training set consists of $17563$ pairs of images (where a  pair consists of a left and right eye image corresponding to a patient). In this dataset, we try and predict from five levels of diabetic retinopathy: no DR ($25810$ images), mild DR ($2443$ images), moderate DR ($5292$ images), severe DR ($873$ images), or proliferative DR ($708$ images). The images are taken in variable conditions: by different cameras,  varying conditions of illumination and different resolutions. These images come from the EyePACS dataset that was used in a Diabetic Retinopathy Detection competition that was hosted on the Kaggle platform. Also, this dataset was used in later works~\cite{de2018weighted, nebot2016diabetic} and ordinal techniques (such as an ordinal cost function) were applied in order to achieve better performance. A validation set is set aside, consisting of $10\%$ of the patients in the training set. The images are resized to 128 by 128 pixels. Data augmentation techniques are applied to achieve a higher number of samples.

@@

	\subsection{The model}
	\label{sect:model}
	A convolutional neural network (CNN) has been used for both datasets. The architecture of this CNN is presented in the Table \ref{table:CNNArchitecture}.

@@

	Every convolutional layer is followed by an ELU activation layer~\cite{clevert2015fast} and a batch normalization~\cite{ioffe2015batch}. At the output, a Proportional Odds Model is used with different link functions~\cite{agresti2010analysis}. The logit link function is commonly used within POM. In this paper, we are comparing other link functions like probit or complementary log-log with the logit link. These three types of links are explained below and represented in Figure \ref{fig:linkfunctions}.

	\begin{itemize}
		\item \textit{Logit}. Logit link function is the most widely used function for Proportional Odds Models. The logit link is shown in \ref{eq:logit}.
		\begin{equation}
		logit[P(y \preceq \mathcal{C}_q)] = b_q - w^Tx, \quad 1, ..., Q-1
		\label{eq:logit}
		\end{equation}
		\item \textit{Probit}. Probit link function is the inverse of the standard normal cummulative distribution function (cdf). Its expression is shown in \ref{eq:probit}.	
		\begin{equation}
		\Phi^{-1}[P(y \preceq \mathcal{C}_q)] = b_q - w^Tx, \quad 1, ..., Q-1
		\label{eq:probit}
		\end{equation}
		\item \textit{Complementary log-log}. Like the logit and the probit transformation, the complementary log-log transformation takes a response that is restricted to the (0,1) interval and converts it into something in $(-\infty, +\infty)$ interval, but unlike the logit and the probit, the complementary log-log function is not symmetric. With a continuous predictor $x$, for example, $P(Y \le j)$ approaches 0 at a different rate than it approaches 1. Complementary log-log expression is shown in  \ref{eq:cloglog}.
		\begin{equation}
		\small
		log[-log[1 - P(y \preceq \mathcal{C}_q)]] = b_q - w^Tx, \quad 1, ..., Q-1
		\label{eq:cloglog}
		\end{equation}
	\end{itemize}

	\begin{figure}[!t]
		\centering
		\includegraphics[width=2.5in]{img/linkfunctions.png}
		\caption{Representation of link functions.}
		\label{fig:linkfunctions}
	\end{figure}

	When the given data is not symmetric in the [0,1] interval and increase slowly at small to moderate value, but increases sharply near 1, the logit and probit models are inappropriate. However, in this situation, the complementary log-log model might give satisfying results.

@@

@@

	The model is evaluated with Quadratic Weighted Kappa metric (QWK)~\cite{ben2008comparison}. This evaluation measure gives a higher weight to the errors that are further from the correct class. Quadratic Weighted Kappa Loss is considered as loss function for the optimizer as it gives better performance for ordinal classification problems.

@@

	\subsection{Factors}
	Three different factors have been considered: learning rate, batch size and link function for the final output layer.

@@

	\begin{table}[!t]
		\caption{Diabetic Retinopathy results. BS stands for Batch Size, LR for Learning Rate and LF for link function.}
		\label{table:DRresults}
		\footnotesize
		\centering
		\begin{tabular}{ccc|cc}
			BS & LR & LF & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			$5$ & $10^{-2}$ & Logit & $0.44888$ & $0.41630$\\
			& & Probit & $0.46724$ & $0.45972$\\
			& & c log-log & $0.42854$ & $0.41448$\\
			& $10^{-3}$ & Logit & $0.56496$ & $0.55356$\\
			& & Probit & $0.57084$ & $0.56392$\\
			& & c log-log & $0.54796$ & $0.53436$\\
			& $10^{-4}$ & Logit & $0.52884$ & $0.52032$\\
			& & Probit & $0.53802$ & $0.52302$\\
			& & c log-log & $0.53824$ & $0.51974$\\
			$10$ & $10^{-2}$ & Logit & $0.54970$ & $0.53142$\\
			& & Probit & $0.52124$ & $0.50806$\\
			& & c log-log & $0.42760$ & $0.42270$\\
			& $10^{-3}$ & Logit & $0.58036$ & $0.57686$\\
			& & Probit & $0.56536$ & $0.55810$\\
			& & c log-log & $\textbf{0.58830}$ & $\textbf{0.58150}$\\
			& $10^{-4}$ & Logit & $0.53158$ & $0.53850$\\
			& & Probit & $0.54560$ & $0.54082$\\
			& & c log-log & $0.53928$ & $0.53742$\\
			$15$ & $10^{-2}$ & Logit & $0.55828$ & $0.55076$\\
			& & Probit & $0.54112$ & $0.53430$\\
			& & c log-log & $0.56676$ & $0.56428$\\
			& $10^{-3}$ & Logit & $0.55748$ & $0.55106$\\
			& & Probit & $0.58182$ & $0.57894$\\
			& & c log-log & $0.56340$ & $0.55938$\\
			& $10^{-4}$ & Logit & $0.54686$ & $0.54342$\\
			& & Probit & $0.53258$ & $0.53278$\\
			& & c log-log & $0.53900$ & $0.53828$
		\end{tabular}
	\end{table}

@@

	\begin{table}[!t]
		\caption{Adience results. BS stands for Batch Size, LR for Learning Rate and LF for link function.}
		\label{table:Adienceresults}
		\footnotesize
		\centering
		\begin{tabular}{ccc|cc}
			BS & LR & LF & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			$5$ & $10^{-2}$ & Logit & $0.72072$ & $0.68408$\\
			& & Probit & $0.77608$ & $0.72574$\\
			& & c log-log & $0.75430$ & $0.70742$\\
			& $10^{-3}$ & Logit & $0.71460$ & $0.67868$\\
			& & Probit & $0.81906$ & $0.77154$\\
			& & c log-log & $0.57196$ & $0.55608$\\
			& $10^{-4}$ & Logit & $0.54432$ & $0.51934$\\
			& & Probit & $0.54360$ & $0.52422$\\
			& & c log-log & $0.54802$ & $0.52484$\\
			$10$ & $10^{-2}$ & Logit & $0.78872$ & $0.72994$\\
			& & Probit & $0.79332$ & $0.74750$\\
			& & c log-log & $0.78004$ & $0.73260$\\
			& $10^{-3}$ & Logit & $0.49686$ & $0.47112$\\
			& & Probit & $0.59778$ & $0.58476$\\
			& & c log-log & $0.52016$ & $0.49968$\\
			& $10^{-4}$ & Logit & $0.60972$ & $0.56814$\\
			& & Probit & $0.54638$ & $0.52632$\\
			& & c log-log & $0.75428$ & $0.52094$\\
			$15$ & $10^{-2}$ & Logit & $0.79104$ & $0.74388$\\
			& & Probit & $\textbf{0.84254}$ & $\textbf{0.78666}$\\
			& & c log-log & $0.78802$ & $0.74624$\\
			& $10^{-3}$ & Logit & $0.56946$ & $0.52978$\\
			& & Probit & $0.70262$ & $0.65416$\\
			& & c log-log & $0.49242$ & $0.47320$\\
			& $10^{-4}$ & Logit & $0.55202$ & $0.52484$\\
			& & Probit & $0.54288$ & $0.51740$\\
			& & c log-log & $0.53826$ & $0.52542$
		\end{tabular}

@@

	\subsection{Statistical analysis}
	\label{sect:statisticalanalysis}
	In this subsection, a statistical analysis will be performed in order to obtain conclusions from the results presented in this section.

@@

	We denote by $WK_{i,j,k,l}(i=1, ..., 3; j = 1, ..., 3; k = 1, ..., 3)$ the value observed when the first factor is at the i-th level, the second at the j-th level and the third at k-th level. We assume that the two factors do not act independently and therefore there exists an interaction between them. In this case, the observations fit \ref{eq:anova3}.

	\begin{equation}
	\begin{aligned}
	\footnotesize
	WK_{i,j,k,l} = & \mu + L_i + P_j + B_k + LP_{i,j} + LB_{i,k} \\& + PB_{j,k} + LPB_{i,j,k} + \epsilon_{i,j,k,l}
	\label{eq:anova3}
	\end{aligned}
	\end{equation}

@@

	\begin{table}[!t]
		\caption{ANOVA III for the analysis of the main factors in the design of a Convolutional Ordinal Neural Network for the Retinopathy dataset.}
		\label{table:ANOVARetinopathy}
		\centering
		\small
		\begin{tabular}{l|ccccc}
			\multicolumn{6}{c}{Response variable WK}\\\hline
			Source & S.S. & D.F. & M.S. & F-ratio & Sig.\\\hline
			Model & $37.860$ & $9$ & $4.207$ & $1562.840$ & $0.000$\\
			$L$ factor & $0.057$ & $2$ & $0.029$ & $10.646$ & $0.000$\\
			$P$ factor & $0.121$ & $2$ & $0.060$ & $22.468$ & $0.000$\\
			$LP$ factors & $0.057$ & $4$ & $0.014$ & $5.261$ & $0.001$\\
			Error & $0.339$ & $126$ & $0.003$ & & \\
			Total & $38.199$ & $135$ & & & 
		\end{tabular}
	\end{table}

@@

	\begin{table}[!t]
		\caption{ANOVA III for the analysis of the main factors in the design of a Convolutional Ordinal Neural Network for the Adience dataset.}
		\label{table:ANOVAAdience}
		\centering
		\small
		\begin{tabular}{l|ccccc}
			\multicolumn{6}{c}{Response variable WK}\\\hline
			Source & S.S. & D.F. & M.S. & F-ratio & Sig.\\\hline
			Model & $52.328$ & $15$ & $3.489$ & $1426.170$ & $0.000$\\
			$L$ factor & $0.027$ & $2$ & $0.014$ & $5.582$ & $0.005$\\
			$P$ factor & $1.031$ & $2$ & $0.516$ & $210.809$ & $0.000$\\
			$B$ factor & $0.089$ & $2$ & $0.045$ & $18.253$ & $0.000$\\
			$LP$ factors & $0.183$ & $4$ & $0.046$ & $18.695$ & $0.000$\\
			$PB$ factors & $0.124$ & $4$ & $0.031$ & $12.695$ & $0.000$\\
			Error & $0.294$ & $120$ & $0.002$ & & \\
			Total & $52.622$ & $135$ & $3.489$ & & 
		\end{tabular}
	\end{table}

@@

	\begin{table}[!t]
		\caption{Nominal method results. BS stands for Batch Size and LR for Learning Rate.}
		\label{table:NominalResults}
		\footnotesize
		\centering
		\begin{tabular}{lcc|cc}
			Dataset & BS & LR & $\kappa_{val}$ & $\kappa_{test}$\\\hline
			Retinopathy & $5$ & $10^{-2}$ & $0.2210$ & $0.2050$\\
			 & & $10^{-3}$ & $0.2807$ & $0.2654$\\
			 & & $10^{-4}$ & $0.4580$ & $0.4449$\\

			 & $10$ & $10^{-2}$ & $0.2972$ & $0.2997$\\
			 &  & $10^{-3}$ & $0.3972$ & $0.3989$\\
			 &  & $10^{-4}$ & $\textbf{0.4972}$ & $\textbf{0.4967}$\\

			 & $15$ & $10^{-2}$ & $0.3692$ & $0.3676$\\
			 & & $10^{-3}$ & $0.4107$ & $0.4162$\\
			 & & $10^{-4}$ & $0.4929$ & $0.4859$\\
			\hline
			Adience & $5$ & $10^{-2}$ & $0.0605$ & $0.0769$\\
			 & & $10^{-3}$ & $0.7994$ & $0.7313$\\
			 & & $10^{-4}$ & $0.8009$ & $0.7383$\\

			 & $10$ & $10^{-2}$ & $0.0559$ & $0.0576$\\
			 & & $10^{-3}$ & $0.8239$ & $0.7436$\\
			 & & $10^{-4}$ & $0.7938$ & $0.7218$\\

			 & $15$ & $10^{-2}$ & $0.0593$ & $0.0713$\\
			 & & $10^{-3}$ & $\textbf{0.8309}$ & $\textbf{0.7575}$\\
			 & & $10^{-4}$ & $0.7867$ & $0.7129$\\
		\end{tabular}
	\end{table}
